### 2023-07-25 SDSS-2023/testing-charts-accuracy.tex
tex file

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{jds}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{lscape}
\author[1]{Kiegan Rice\thanks{Corresponding author. Email: rice-kiegan@norc.org}}
\author[2]{Heike Hofmann\footnote{Email: hofmann@iastate.edu}}
\author[1]{Nola du Toit\footnote{Email: dutoit-nola@norc.org}}
\author[1]{Edward Mulrow\footnote{Email: mulrow-edward@norc.org}}
\affil[1]{NORC at the University of Chicago}
\affil[2]{Department of Statistics, Iowa State University}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Testing Charts: viewers' perceptual accuracy in surveys},
  pdfkeywords={list, keywords, here, 3-6, are, fine},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Testing Charts: viewers' perceptual accuracy in surveys}
\author{}
\date{}

\begin{document}
\maketitle
\begin{abstract}
The use of visuals is a key component in scientific communication, and
decisions about the design of a data visualization should be informed by
what design elements best support the audience's ability to perceive and
understand the components of the data visualization. We build on the
foundations of Cleveland and McGill's work in graphical perception,
employing a large, nationally-representative, probability-based panel of
survey respondents to test perception in statistical charts. Our
findings provide actionable guidance for data visualization
practitioners to employ in their work.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, breakable, enhanced]}{\end{tcolorbox}}\fi

  \newcommand{\hh}[1]{{\textcolor{orange}{#1}}} 
  \newcommand{\kr}[1]{{\textcolor{teal}{#1}}}.  
  \newcommand{\ejm}[1]{{\textcolor{ForestGreen}{#1}}}.  
  \newcommand{\ndt}[1]{{\textcolor{purple}{#1}}}.

% take out everything above later
% keep everything below

  \newcommand{\blandscape}{\begin{landscape}} 
  \newcommand{\elandscape}{\end{landscape}}

  \setlength{\parindent}{0pt}
  \singlespacing

\ndt{break up first sentence}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

What do viewers see when we show them a data chart? A data chart -- at
its core -- maps values to graphical elements: quantitative elements are
represented as measurable features, such as position, size, or shade.
\hh{do we need to also go into qualitative variables?} Modern data
visualizations are much more than a simple, objective mapping of values
to a plane; they contain contextual and design elements, and are often
structured to support the viewer in understanding a particular view of a
set of data or specific pattern underlying the values. The design of a
data visualization impacts a viewer's ability to achieve that
understanding; a poorly designed data visualization may leave viewers
struggling to understand the content or context, or make it difficult to
complete accurate and useful comparisons of values across groups or time
points. More broadly, the design of a data visualization can change how
viewers interact with the chart.

\ndt{define structural}

\ndt{The viewer's employment of comparisons of the components is a crucial step in the process of interacting with and understanding of chart.}

Cleveland and McGill (1984) observed as such, and in their seminal study
defined the better visual among a pair as the one that allows viewers to
make more accurate comparisons. Based on mappings of quantitative
variables to different graphical elements
\ndt{example of element to explain it XXX}, Cleveland and McGill's study
resulted in a ranking of perceptual tasks \ndt{give example of task}
from most accurate to least accurate, which was then extended by
Mackinlay (1986) to a theoretical framework ranking tasks' order along
their ordinal and nominal scales.

Cleveland and McGill's work -- while a foundational user study in
graphical perception -- utilized a small convenience sample, consisting
of only a few individuals recruited from among the authors' coworkers
and their spouses. Heer and Bostock (2010) reproduced Cleveland and
McGill's rankings using a larger sample from a crowd sourcing platform,
employing a total of 82 Amazon Mechanical Turk (mTurk) workers for the
study. Crowd sourced samples were shown by Borgo et al. (2017) to be
biased towards more male, younger, and relatively higher education
relative to the adult U.S. population as a whole; generally, convenience
and opt-in samples are not representative of the general population, a
common target audience for data visualization and scientific
communication work. Further, the populations' emphasis on higher
education individuals also leads to results which hold for groups of
individuals who may be more likely to have prior exposure to data
visualization in the context of scientific communication, or more
exposure to data topics in higher education, but may not hold across
other groups within the population.

Here, we seek to answer whether it is possible to reproduce some of the
previous findings in the context of a survey with a large,
nationally-representative set of respondents, and within that context we
focus on the following research questions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  How do structural design choices in a data visualization impact
  viewers' ability to identify the larger of two elements?
\item
  How is viewer interaction with the task impacted by structural design
  choices in a data visualization?
\item
  Are there differences in perception and interaction with the tasks
  across demographic groups?
\end{enumerate}

We employ a probability-based survey panel and run a series of
perception tests with nationally-representative samples of respondents
from that panel. The advantage of using a probability-based approach is
two-fold. First, we have access to a large sample of survey participants
and thus have greater power in making inference about graphical
perceptional abilities. Second, the sample is representative of the
general adult public in the U.S., and this allows us to test whether
prior results from convenience samples hold with a nationally
representative sample and whether there are differences in those results
across demographic subgroups.

\ndt{I think the next paragraph can be moved above the RQs. Use this paragraph to set up the concepts, define constructs, then explain the study, and then note the RQs.}

Specifically, we present viewers with structural variations of bar
charts and ask them to answer questions comparing the size of elements
within those charts. In this work, we present the series of tests we
completed and the resulting findings. The remainder of the paper is
organized as follows: first, we describe the design of the visual
stimuli used in the perception tests. We then describe the population of
study respondents and obtained survey sample. Subsequently, we share a
summary of the resulting survey responses across each of the tests,
including analyses on accuracy of responses and response behavior.
Finally, we discuss implications of this work and next steps.

\hypertarget{study-design-stimulus}{%
\subsection{Study design -- stimulus}\label{study-design-stimulus}}

\ndt{
I think this paragraph should go first in this section.
1. the survey format
2. the type of data
3. The stimulus
3. how we set up the stimulus and why
4. the tasks and alighed v unaligned
4. qtext
5. fig 1}

Each task is made up of two elements: a visual stimulus and a question
about the stimulus. In our study, each visual stimulus is an image of a
data visualization, while each question prompts viewers to identify
which of two marked pieces in the data visualization is larger.

The comparison between sets of marked pairs is intentionally designed to
be a difficult task, with the difference between the values represented
in the two marked pieces being chosen close to their just-noticeable
difference. The \textbf{Just-Noticeable Difference} (JND) is defined as
the smallest difference that will be detected 50\% of the time. Prior
results from studies on bar charts and pie charts (Lu et al. 2022)
inform the differences in charts shown to survey panelists.

We employ comparisons at the JND in our tasks in order to maximize our
ability to identify the impact of design changes on viewer accuracy and
behavior. Asking perception tasks in a survey differs from the
controlled environment of a cognitive lab, where these kind of questions
may usually be assessed. Rather than asking the same (or similar) type
of question with varied signal strength dozens or hundreds of times, we
are limited to only a few questions at a time.
\ndt{note why we are limited to only a few questions at a time - that is not clear.}
With a small set of tasks, we need to present tasks that are
perceptually hard, and thus ask questions about stimuli that are close
to our perceptual threshold. Therefore, we focus on questions which vary
the structure of the presented image, but ask viewers to compare the
same underlying data values across those varied images.

We ask participants to determine which of two just-noticeably different
marked pieces \ndt{is pieces the same as tiles below?} is larger within
each chart, and focus on three main sets of structural variation in the
design of that chart. First, we vary the alignment of the pieces in
question. \ndt{perhaps a bulleted list would work better here} Viewers
are presented with two marked pieces in a chart that do not share a
common baseline, then two pieces that do share a common baseline.
\ndt{"we refer to this as aligned and unaligned"} Second, we vary the
orientation of the chart -- we rotate the vertical stacked bar chart,
and present a horizontally oriented version of the same chart, with
identically sized marked pieces. Finally, we change the aspect ratio of
the chart and present a wider version of the horizontally oriented chart
which has longer, but thinner, marked pieces.

\hh{Not sure where should put this, parking it here for right now:}
\hh{expectations - start}

What we call `aligned' and `unaligned', here, is similar to Cleveland
and McGill's positions along aligned and unaligned scales, but with some
modifications. Figure~\ref{fig-tasks-explain} gives an overview of the
comparisons of tasks 1 through 3 and the closest corresponding tasks in
Cleveland and McGill.

In this paper, both `aligned' and `unaligned' bars share the same axis.

Aligned tiles are additionally anchored in the same position along the
axis, i.e.~the difference between their sizes can be reduced to a
positional assessment, which in the absence of three-dimensional cues
seems to be the general strategy viewers use when evaluating barcharts
(Zacks et al. 1998).\\
Unaligned tiles do not share this anchor of position, i.e.~the
difference between their sizes needs at least two comparisons of
positions. However, unaligned bars in a stacked barchart are not just
assessments of the lengths of bars -- the other tiles in the chart
provide a reference frame of the shared axis, which \emph{should} help
with an assessment of the tiles' sizes. As shown in the sketch of
Figure~\ref{fig-tasks-explain}, this framing leads to a stimulus that is
between Cleveland and McGill's task of `position along the same but
unaligned axis' and an assessment of length.

We would expect that comparing unaligned tiles is a harder task (with
correspondingly lower levels of accuracy) than a comparison of aligned
tiles, with the framing given by the other tiles in the same column
mitigating some of this difficulty.
\hh{In particular, the additional context of the other tiles should *help* rather than *hurt* the comparison, unlike other situations, where additional information besides the pieces involved in the comparison were found to be distractors}
(Talbot, Setlur, and Anand 2014).

\begin{figure*}[hbt]

{\centering \includegraphics[width=0.6\textwidth,height=\textheight]{images/sketch-tasks.png}

}

\caption{\label{fig-tasks-explain}Comparisons made in charts within the
Cleveland and McGill ranking}

\end{figure*}

\ndt{"The survey format guides the design and structure of the questions asked...." Why is this so?}

Our use of a survey format guides the format and design of the questions
asked and how they are presented to respondents. First, participant
instructions must be delivered in a very short and easily understandable
format, because participants cannot ask clarifying questions about the
task as they might be able to in a cognitive lab setting. Second, we
want to utilize content within the data visualization image which is not
socially or politically charged for the average U.S. adult; this risks
participants reacting to the subject matter within the chart rather than
focusing on the presented task. \ndt{stimulus?} For this reason, we
utilize data on living arrangements of older U.S. adults -- a topic
which most U.S. adults will have some familiarity with, but is not
inherently polarizing. Finally, to prevent viewers from being exposed to
slight variations of the same stimulus in a row (and risk unforeseen
order effects or respondents using prior questions to inform their
responses), we either split a survey sample in two and show each
subsample a distinct structural version of the chart or test variations
of a chart across distinct rounds of the survey.

\hypertarget{question-text}{%
\paragraph{Question text}\label{question-text}}

When viewing each chart, participants were asked to compare the relative
sizes of marked elements within the chart:

\ndt{This should go above fig 1 to explain the set up for fig 1}

``There are many charts used in the news media to portray data visually.
Looking at the chart below, which of the marked dark blue pieces is
bigger, A or B? Just your best guess is fine.

\begin{itemize}
\item
  A is bigger
\item
  B is bigger
\item
  They are the same
\end{itemize}

When presented with the aligned version of each chart, pieces were
marked with a C and D and the question text is updated accordingly. In
all scenarios, the second option (B {[}D{]} is bigger) is the correct
response.

For a given task, viewers are first presented with the unaligned version
of the task, followed by the aligned version of the task. The time in
seconds that each respondent spent on each task was recorded, as well as
whether the participant zoomed in on each chart while answering the
question. Respondents were also asked to rate their certainty in their
response to each question on a five-point scale.

\begin{figure*}[hbt]

{\centering \includegraphics[width=0.9\textwidth,height=\textheight]{images/stimuli-overview.png}

}

\caption{\label{fig-tasks}Each of the visual stimuli presented to
viewers. In each bar chart, two pieces are marked. The larger piece in
each chart are pieces B and D.}

\end{figure*}

\hypertarget{task-1-vertical-stacked-bar}{%
\subsubsection{Task 1: Vertical stacked
bar}\label{task-1-vertical-stacked-bar}}

\autoref{fig-tasks}(a,d) shows the two stacked bar charts shown to
participants in the first task. The marked tiles in each plot are 155
pixels apart, which leads to a JND of 3.5 pixels based on Lu et al.
(2022)`s model. The heights of the bars are 205 (left) and 213 pixels
(right), corresponding to about twice the JND. This difference should
lead to a relatively high accuracy rate for participants and
simultaneously limit the amount of frustration resulting from a task
that is perceived as 'too hard'.

Both charts show the same data with slight modifications to the order of
the levels -- the first and second level in each of the bars are
reversed between the left and the right chart. Both charts are displayed
at the same size, i.e.~in both cases both the difference in size between
the bars and the horizontal distance between the bars is the exact same
amount. This leaves the vertical positioning of the bars as the only
difference between the charts. Any differences in observed responses can
therefore be attributed to this difference in presentation.

Task 1 was presented to viewers in two rounds (Rounds 1-2), with four
color scheme variations each presented to 50\% of respondents in each
round. There were no significant differences in respondent accuracy or
behavior across color schemes, and we combine them here into one set of
stimulus responses to focus on the comparison of structural differences.
\kr{Should we put the color variations and a test for significance in accuracy/a visual in the appendix/supplementary materials? }
\hh{we could do that, but I'm not sure that we need to mention color at all.}

\hypertarget{task-2-horizontal-stacked-bar}{%
\subsubsection{Task 2: Horizontal stacked
bar}\label{task-2-horizontal-stacked-bar}}

The visual mappings in the second task, shown in \autoref{fig-tasks}(b,
e), are identical to the first task, but the axes of the chart are
rotated so that the stacked bars are represented in a horizontal format.
This represents a structural change in how the data are presented to the
viewer while preserving the pixel size of the elements viewers must
compare. Tasks 2 and 3 were asked within the same survey round (Round 3)
and the full sample was randomly split among respondents, with 50\% of
participants seeing the Task 2 stimuli and 50\% of participants seeing
the Task 3 stimuli.

\hypertarget{task-3-horizontal-wide-stacked-bar}{%
\subsubsection{Task 3: Horizontal wide stacked
bar}\label{task-3-horizontal-wide-stacked-bar}}

The images utilized in the third task again represent the same data as
in the first two tasks, and the overall image has dimensions which are
identical to the images in the first two tasks, displaying at the same
size during the survey.
\hh{The length of bars is increased to fit the new dimensions. This increases the difference in the length of bars from previously 8 pixels to 13 pixels. The widths of the tiles are adjusted accordingly (from 50 to 30 pixels) to keep the overall area of the tiles approximately constant. }
\ndt{agree to drop note on color. Can perhaps just say we used CBlind  friendly}

\hypertarget{task-3b-floating-bars}{%
\subsubsection{Task 3b: Floating bars}\label{task-3b-floating-bars}}

\hh{
In order to measure the impact of the chart's context, we asked a subset of participants to assess the relative sizes of the two tiles A and B (as shown in the bottom right of @fig-tasks-explain) without giving any additional information.}

\hypertarget{study-design-participants}{%
\subsection{Study design --
participants}\label{study-design-participants}}

Participants were recruited as part of NORC's AmeriSpeak panel, which
utilizes a probability-based sampling methodology and samples U.S.
households from NORC's National Sample Frame that provides coverage of
over 97\% of U.S. households. The current panel size is 54,001 panel
members aged 13 and over residing in over 43,000 households (Dennis
2019, updated 2022). \textbackslash ejm\{46,565 panelists are CAWI; they
are 93.3\% of the panel and 93.6\% of the total panel weights.\} Each
test was conducted using the AmeriSpeak Omnibus survey, which runs
biweekly and samples around 1,000 U.S. adults to answer questions on a
variety of topics. Our tests require visual inspection of an image; for
this reason, our survey questions were only presented to web-based
panelists and not panelists who respond via phone interviews.

The Omnibus sample is not longitudinal in nature, i.e.~we do not have
any information about whether the same panelists were included in
multiple rounds. While there is a (small) chance that panelists are
included in multiple rounds of our data, there was at least a one month
gap between viewing each visual stimulus, and for the purposes of the
analysis we will consider data from these viewings as independent.

\hypertarget{study-design-survey-weighting}{%
\subsection{Study design -- survey
weighting}\label{study-design-survey-weighting}}

Paragraph on how survey weights are derived? -- \textbf{Question for Ed
-- is this something AmeriSpeak can provide boilerplate language for?}

All calculations in this paper are done in R (R Core Team 2022), and
weights are applied in analyses using the \texttt{survey} package
(Lumley 2004) version 4.0 (Lumley 2020) based on Lumley (2010).

When combining responses from different surveys, weights
\ejm{are rescaled} with respect to the total population, so that their
order after combining still reflects their relative importance.

We \emph{combine} (rather than cumulate) a set of \(\ell\) surveys
\(S_1\), \(S_2\), \ldots{} \(S_\ell\) (with \$ \ell \ge 2\$ ) as
described in O'Muircheartaigh and Pedlow (2002), by multiplying weights
in \(S_i\) by \(\lambda_i\), for \(1 \le i \le \ell\).
\(\lambda_i \in [0,1]\) with \(\sum_i \lambda_i = 1\) is given as \[
\lambda_i = \frac{n_i/d_i}{\sum_{j=1}^{\ell}n_j/d_j},
\] where \(n_i\) is the nominal sample of survey \(S_i\) and \(d_i\) are
the design effects for the estimators. Here, \(d_i\) are estimated as \[
d_i = 1 + CV(w \in S_i)^2
\] where \(CV\) is the coefficient of variation of the weights \(w\)
within each sample, and is estimated as in Kish (1965) : \[ 
CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
\]

The data for this paper were collected in several rounds as part of the
NORC Omnibus. The resulting number of participants, effective sample
sizes, and \(\lambda\) values are shown in \autoref{tbl-rounds}.

\hypertarget{tbl-rounds}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}@{}}
\caption{\label{tbl-rounds}Survey rounds: dates, number of participants
(nominal sample size), effective sample size, sum of weights, and
factors for the combining surveys as discussed above.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Date
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\# Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Effective Sample Size
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum of weights \(\sum_i w_i\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\lambda_i\)
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Date
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\# Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Effective Sample Size
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum of weights \(\sum_i w_i\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\lambda_i\)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Round 1 & April 2022 & 933 & 521.1 & 934.9 & \(\lambda_1\) = 0.343 \\
Round 2 & May 2022 & 953 & 485.7 & 953.4 & \(\lambda_2\) = 0.320 \\
Round 3 & June 2022 & 921 & 513.5 & 923.1 & \(\lambda_3\) = 0.338 \\
\textbf{Total} & --- & 2807 & 1520.8 & & \\
\end{longtable}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{respondents}{%
\subsubsection{Respondents}\label{respondents}}

A total of 2807 respondents participated across the three rounds. The
number of responses and corresponding effective sample sizes in each
round are shown in \autoref{tbl-rounds}. All responses were combined
into one set of survey responses, with adjusted combined sample weights
and indicators for which task each respondent was exposed to. The
resulting sample sizes corresponding to each task are shown in
\autoref{tbl-tasks}.

\hypertarget{tbl-tasks}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}@{}}
\caption{\label{tbl-tasks}Survey tasks: number of participants (nominal
sample size) and effective sample size for each task.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\# Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Effective Sample Size
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\# Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Effective Sample Size
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Task 1 & Vertical & 1886 & 1007.1 \\
Task 2 & Horizontal & 459 & 266.5 \\
Task 3 & Horizontal wide & 462 & 249.1 \\
Task 3b & Floating bars & 915 & 473.0 \\
\end{longtable}

Distributions of demographic characteristics across each task are shown
in \autoref{fig-rounds-demographics}, the dark points show percentages
for US adults based on the 5 year estimates of the American Community
Survey 2021. The distribution of demographics are all quite similar
across tasks and their 90\% margin of errors (hashed lines) for the most
part cover the Census estimates.

\begin{figure*}[hbt]

{\centering \includegraphics{./figures/fig-rounds-demographics-1.png}

}

\caption{\label{fig-rounds-demographics}Demographics of respondents in
each of the tasks. The dots show percentages of the national
demographics as estimated by the American Community Survery (ACS) 2021.}

\end{figure*}

\hypertarget{accuracy-of-responses}{%
\subsubsection{Accuracy of responses}\label{accuracy-of-responses}}

We first investigate respondent accuracy in selecting the correct
response. As participants were able to select the option `They are the
same', there are several ways to model accuracy and response. The
argument could be made that for the purposes of practical
interpretation, `they are the same' is a correct choice, as the options
are visually very similar and not substantially different values within
the context of the data shown in the chart. However, we are interested
in understanding whether viewers \emph{can} perceive the difference and
correctly identify which piece is larger and thus consider multiple ways
of modeling response to investigate participant response patterns. We
begin by defining a measure of binary `correctness', for which all
answers that are not the correct option (B {[}D{]} is bigger) are
`incorrect', including the selection of `they are the same'.

\autoref{fig-alignment}(a) displays all responses along the binary
correctness measure, separated by whether the stimulus was an aligned
task or unaligned task. We can see that levels of accuracy for all
responses are significantly higher for the `easier' (aligned) task, with
about twice as many respondents correctly selecting the larger of the
two marked elements.

Because each participant was shown both the aligned and unaligned
versions of the chart, we can use a paired \(t\)-test to compare mean
accuracy between the two charts. The resulting \(t\)-statistic is highly
significant (\(t\) statistic: 21.2, df: 2265, \(p\)-value:
\(\ll 0.0001\)).

\begin{figure*}[hbt]

{\centering \includegraphics[width=0.9\textwidth,height=\textheight]{./figures/fig-alignment-1.png}

}

\caption{\label{fig-alignment}On the left (a), a stacked barchart shows
the number of respondents with correct (green) and incorrect (grey)
responses to the two comparison questions. When tiles are aligned along
the same axis, more than twice the number of responses are correct. The
shaded area along the top of the green tiles corresponds to 95\%
confidence intervals around (marginal) correct responses. On the right
(b), a parallel coordinate plot shows all combinations of responses.
There's a huge asymmetry in the number of responses where participants
answered only one of the questions correctly. A lot more responses are
correct when comparing aligned tiles than unaligned tiles.}

\end{figure*}

Next, we consider an ordinal model to investigate response behavior
across all three options -- `A {[}C{]} is bigger', `B {[}D{]} is
bigger', and `They are the same', and consider these response patterns
across each of the stimuli.

\begin{figure*}[hbt]

{\centering \includegraphics[width=0.9\textwidth,height=\textheight]{./figures/fig-response-123-1.png}

}

\caption{\label{fig-response-123}Responses for accuracy by task across
the designs. The overlaid rectangles represent 95\% confidence
intervals. The letters in blue and orange encode significances between
pairwise proportions: two bars have a significantly different proportion
(at a 5\% significance level) if they do not share a letter. There is no
significant difference between the designs for wrong responses. When
tiles are unaligned, the horizontal wide barchart is showing the highest
accuracy. For aligned tile, the horizontal wide design and the vertical
design do not show a significant difference in accuracy. Floating bars
have a significantly lower rate of correct answers than all the other
designs.}

\end{figure*}

Figure~\ref{fig-response-123} shows the results of a cell-means model
with ordinal response \(Y_k\), where \(Y_k\) is the \(k\)th
participant's response, \(Y_k \in \{1, 2, 3\}\), where `correct' is
encoded as 1, `they are the same' is encoded as 2, and `wrong' is
encoded as 3:

\[
\text{logit }P(Y_k \le \ell) = \mu_{ij\ell(k)},
\]

where \(\ell \in \{1, 2\}\); \(i \in \{1, 2\}\) is the comparison type
(1 = Aligned, 2 = Unaligned), and \(j \in \{1, 2, 3\}\) is the chart
design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide. The
estimated values and 95\% confidence intervals are shown in
Table~\ref{tbl-rounds-123}. The lowercase letters indicate significances
between pairs of estimates: two estimates (in the same column) are
significantly different at 5\% level, if they do not have any letter in
common (Piepho 2004).

\hypertarget{tbl-rounds-123}{}
\setlength{\LTpost}{0mm}
\begin{longtable}{lrrlrrl}
\caption{\label{tbl-rounds-123}Odds for the cell-means model, letters behind numbers indicate pairwise
significances. Within the same \emph{column} values are significantly
different (at 5\%) if they do not share the same letter. }\tabularnewline

\caption*{
{\large Odds of accuracy by task and chart type}
} \\ 
\toprule
 & \multicolumn{3}{c}{correct | same or wrong} & \multicolumn{3}{c}{correct or same | wrong} \\ 
\cmidrule(lr){2-4} \cmidrule(lr){5-7}
 & Est. &           [95\% CI] &   & Est. &           [95\% CI] &   \\ 
\midrule
\multicolumn{7}{l}{Unaligned} \\ 
\midrule
Floating bars\textsuperscript{\textit{1}} & $0.08$ &  [$0.06$, $0.12$] & a & $12.35$ &  [$8.10$, $18.83$] & ab \\ 
Horizontal & $0.22$ &  [$0.15$, $0.32$] & b & $5.54$ &  [$4.04$, $7.59$] & c \\ 
Horizontal wide & $0.26$ &  [$0.19$, $0.37$] & bc & $6.20$ &  [$4.46$, $8.61$] & ac \\ 
Vertical & $0.41$ &  [$0.36$, $0.47$] & cd & $6.90$ &  [$5.75$, $8.28$] & ac \\ 
\midrule
\multicolumn{7}{l}{Aligned} \\ 
\midrule
Horizontal & $0.63$ &  [$0.49$, $0.81$] & d & $14.02$ &  [$9.23$, $21.30$] & b \\ 
Horizontal wide & $2.67$ &  [$2.04$, $3.48$] & e & $17.12$ &  [$10.49$, $27.94$] & b \\ 
Vertical & $1.51$ &  [$1.33$, $1.71$] & f & $11.33$ &  [$8.99$, $14.28$] & b \\ 
\bottomrule
\end{longtable}
\begin{minipage}{\linewidth}
\textsuperscript{\textit{1}}We are including here, responses on an A vs B comparison for floating bars, corresponding to a length assessment, see bottom right of Figure 1.\\
\end{minipage}

One pattern in accuracy holds across each of the three structural
variations: \textbf{the aligned task has a higher level of accuracy than
its unaligned counterpart}. Interestingly, while we expect an
improvement in accuracy when shifting from the horizontal to the
horizontal wide design given the larger difference in pixel length
between the two pieces, the resulting effects on the accuracy of the
responses are not completely straightforward: the shift from a vertical
to the (tall) horizontal design is detrimental to an accurate perception
for both aligned and unaligned comparisons. The re-scaled design of the
wide horizontal bars reclaims some of the loss for unaligned bars and
outperforms the vertical design by a similar margin in aligned bars, but
does not out-perform the vertical design when comparing unaligned tiles.
\hh{The beneficial effect of frames/context of the chart is apparent when comparing accuracy of responses for the Horizontal wide design and the floating bars: the rate of correct responses in floating bars is significantly lower than in the wide horizontal charts. This shows that the additional context }

Rates of selecting the incorrect response are low across all three
structural variations and the aligned and unaligned tasks; while viewers
select the incorrect response at significantly higher rates for all
three unaligned tasks relative to the aligned tasks, those rates do not
differ significantly across the three structural variations. Most of the
observed differences in response patterns across structural variations
are attributed to respondents' selection between the correct option or
the `they are the same' option.

To better comprehend these observed patterns, we investigate viewer
interaction with the tasks more broadly as well as differences in
response selection across demographic groups.

\hypertarget{respondent-behavior}{%
\subsubsection{Respondent behavior}\label{respondent-behavior}}

One contributing factor to the observed patterns in response accuracy
might be the way that participants interact with the different designs.
Across all tasks, about half of all participants make use of the option
to zoom into charts. We observe that while zooming does help with the
overall accuracy (which is in agreement with the findings by Lu et al.
(2022) about the physical size of stimuli), the increase is not
significant. However, different designs lead to different rates of
zooming: we observe in \autoref{fig-zoom-123} that when dealing with the
vertical design, the rate of zooming is significantly higher than for
the two horizontal designs.

\begin{figure*}[hbt]

{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{./figures/fig-zoom-123-1.png}

}

\caption{\label{fig-zoom-123}Zooming - not significant for accuracy or
task, but changes by the type of chart.}

\end{figure*}

To formalize this pattern in a model, let \(Y_{jk}\) describe the
zooming behavior of panelist \(k\) on task \(j\). We model zooming
behavior (no = 0, yes = 1) as a logistic regression by correctness of
response (\(\rho\)), task (\(\tau\)), and design (\(\delta\)) of the
chart:

\[
\text{logit } P(Y_k \le 1) = \mu + \rho_{i(k)} + \tau_{j(k)} + \delta_{\ell(k)},
\]

The resulting model estimates and confidence intervals are displayed in
\autoref{tbl-zoom-123}. We observe that zooming behavior \emph{only}
differs significantly by structural design; rates do not differ
significantly between the aligned and unaligned tasks, nor do they
differ significantly by the chosen response. This higher rate of zooming
does not necessarily lead to higher accuracy; although rates of accuracy
are significantly higher for the vertical orientation relative to the
horizontal orientation, they are not significantly higher -- and in
fact, are significantly lower on the aligned task -- than the accuracy
for the horizontal wide orientation.
\hh{I'll pretty up @tbl-zoom-123 if we are going to keep it.}
\kr{I think we should! it's nice to have a model for this}\hh{gt and rmarkdown are frustrating! I'll re-set the table in its final form directly in tex}

\hypertarget{tbl-zoom-123}{}
\begin{longtable}{lrrrr}
\caption{\label{tbl-zoom-123}Coefficients for logistic regression of zooming by task }\tabularnewline

\caption*{
{\large Estimates for logistic regression on zooming behavior}
} \\ 
\toprule
term & estimate & SE & t-statistic & p-value \\ 
\midrule
\multicolumn{5}{l}{} \\ 
\midrule
\$\$\textbackslash{}hat\{\textbackslash{}mu\}\$\$ & $-0.35$ & $0.15$ & $-2.4$ & 0.0155 \\ 
\midrule
\multicolumn{5}{l}{Response} \\ 
\midrule
\$\$\textbackslash{}text\{They are the same \} \textbackslash{}widehat\{\textbackslash{}rho\}\_\{2\}\$\$ & $-0.17$ & $0.09$ & $-1.9$ & 0.0578 \\ 
Wrong \$\$\textbackslash{}widehat\{\textbackslash{}rho\}\_\{3\}\$\$ & $-0.06$ & $0.13$ & $-0.5$ & 0.6466 \\ 
\midrule
\multicolumn{5}{l}{Task} \\ 
\midrule
Aligned \$\$\textbackslash{}widehat\{\textbackslash{}tau\}\_\{2\}\$\$ & $0.00$ & $0.06$ & $-0.1$ & 0.9435 \\ 
\midrule
\multicolumn{5}{l}{Chart design} \\ 
\midrule
Horizontal wide \$\$\textbackslash{}widehat\{\textbackslash{}delta\}\_\{2\}\$\$ & $0.27$ & $0.15$ & $1.8$ & 0.0752 \\ 
Vertical \$\$\textbackslash{}widehat\{\textbackslash{}delta\}\_\{3\}\$\$ & $0.98$ & $0.13$ & $7.7$ & < 0.0001 \\ 
\bottomrule
\end{longtable}

\hypertarget{how-certain-are-participants}{%
\subsubsection{How certain are
participants?}\label{how-certain-are-participants}}

Using linear scores for the response of \texttt{Certainty}, with
\texttt{not\ certain\ at\ all} assigned a score of 1 and
\texttt{extremely\ certain} assigned a score of 5, we can estimate the
effects of task, chart design, and correctness on certainty by using a
cell-means model of the form: \[
Y_{k} = \mu_{ij\ell(k)} + \epsilon_{k},
\] where \(k = 1, ..., N\), \(\mu_{ij\ell(k)}\) is average certainty
(measured on a scale from 1 to 5) of the four combinations of task and
correctness by each of the three designs, where \(i = 1, 2\) encodes
unaligned/aligned, and \(j=1,2\) encodes wrong, correct, and
\(\ell = 1, 2, 3\) encodes horizontal, wide horizontal, and vertical
stacked bar charts, respectively. We also assume that errors are
normally distributed,
i.e.~\(\epsilon_k \stackrel{i.i.d}{\sim} N(0, \sigma^2)\) for all
\(k = 1, ..., N\). The results are shown in Figure~\ref{fig-certainty}.
What we find, is that the highest scores for certainty are associated
with comparing aligned tiles. Certainty scores are not significantly
different between correct and wrong responses. Scores are (mostly)
significantly lower for the unaligned task. Interestingly, the lowest
scores of certainty are associated with correct responses, here.

\begin{figure*}[hbt]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{./figures/fig-certainty-1.png}

}

\caption{\label{fig-certainty}Certainty by task, chart design, and
correctness. Correct answers in the unaligned task have the lowest
certainty scores associated with them. Incorrect responses on the
unaligned task show a significant boost in certainty in the vertical
stacked bar. For horizontal stacked bars the difference between aligned
and unaligned charts matters significantly. Errorbars around the points
show estimated scores (on a scale of 1 to 5) with corresponding 95\% CI.
The letters underneath the bars indicate significances. The scores for
two bars are significantly different if they do not share a letter.}

\end{figure*}

\hypertarget{differences-across-demographic-groups}{%
\subsubsection{Differences across demographic
groups}\label{differences-across-demographic-groups}}

Finally, we turn to investigating responses across demographic groups, a
core benefit provided by the large and representative nature of our
survey samples. \autoref{fig-demographics} displays response patterns
across our three structural variations by gender, age (4 groups),
educational attainment (5 groups), and income level (4 groups). Our
pattern of higher accuracy in selecting the correct response on the
aligned task relative to the unaligned task holds across all demographic
groups and structural variations. However, we do observe different
response patterns, particularly on the unaligned task, across
demographic groups.

Let \(Y_k\) be the response of participant \(k\), on a scale from 1 =
`wrong', 2 = `they are the same' to 3 = `correct'. We use a generalized
cumulative logistic regression, where \(\mu_i\) are intercepts
\(1 \le i < 3\), \(X_k\) are demographics of the \(k\)th participant (in
form of the model matrix), and \(\beta_{i}\) are the coefficients.

\[
\text{ logit } P(Y_k \le i \mid X_k) = \mu_i + X_{k}'\beta_{i}
\] The resulting model estimates and confidence intervals are shown in
\autoref{tbl-demographics}. When considering the easier (aligned) task,
response patterns do not differ significantly by age, gender, or
education level. We do observe, however, a significantly higher log odds
of selecting the `correct' or `they are the same' response (or
significantly lower log odds of selecting `incorrect') among those with
an income between \$60,000 and \$100,000. When we consider the more
difficult (unaligned) task, we see significant separation in response
patterns across gender, educational attainment, and income groups. We do
not see significant differences in responses by age for either task.
Interestingly, we see lower log odds of selecting the `correct' response
among those with higher educational attainment (those with a bachelor's
degree or post graduate study) \emph{and} lower log odds of selecting
the `incorrect' response; as income and educational attainment increase,
respondents are more likely to select the `they are the same' option
during the difficult task. This speaks to the difficulty of the task,
and respondents' interpretation of the task when it is more difficult to
perceive small differences.

\begin{figure}

\begin{minipage}[t]{\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{./figures/fig-demographics-1.png}

}

}

\subcaption{\label{fig-demographics-1}Comparisons of aligned tiles}
\end{minipage}%
\newline
\begin{minipage}[t]{\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{./figures/fig-demographics-2.png}

}

}

\subcaption{\label{fig-demographics-2}Comparisons of unaligned tiles}
\end{minipage}%

\caption{\label{fig-demographics}Panelists' demographics matter,
particularly, when the task difficulty increases. For aligned tiles,
gender, age, and education are not significant factors. However, income
levels do have a (small) effect. When income levels increase, the
percentage of wrong answers (orange) decreases, while the percentage of
correct answers (blue) increases slightly (significant at below 0.05).
For the more difficult task of comparing unaligned tiles, demographics
are more significant. Higher levels of education and higher levels of
income are associated with a significant increase of panelists choosing
a response of `they are the same', resulting in significant decreases of
both correct answers and wrong answers.}

\end{figure}

\blandscape

\hypertarget{tbl-demographics}{}
\setlength{\LTpost}{0mm}
\begin{longtable}{lrrlrrlrrlrrl}
\caption{\label{tbl-demographics}Demographics matter for perception, particularly when the tasks get
harder. }\tabularnewline

\caption*{
{\large Odds of accuracy by task and demographics of respondents}
} \\ 
\toprule
 & \multicolumn{6}{c}{Aligned tiles} & \multicolumn{6}{c}{Unaligned tiles} \\ 
\cmidrule(lr){2-7} \cmidrule(lr){8-13}
 & \multicolumn{3}{c}{correct | same or wrong} & \multicolumn{3}{c}{correct or same | wrong} & \multicolumn{3}{c}{correct | same or wrong } & \multicolumn{3}{c}{correct or same | wrong } \\ 
\cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10} \cmidrule(lr){11-13}
  & Est. &           [95\% CI] &   & Est. &           [95\% CI] &   & Est. &           [95\% CI] &   & Est. &           [95\% CI] &   \\ 
\midrule
\multicolumn{13}{l}{Intercept} \\ 
\midrule
 & $1.51$ &  [$0.91$, $2.50$] &  & $12.75$ &  [$5.59$, $29.10$] & *** & $0.60$ &  [$0.35$, $1.03$] & . & $3.24$ &  [$1.70$, $6.19$] & *** \\ 
\midrule
\multicolumn{13}{l}{Gender} \\ 
\midrule
Female & $0.82$ &  [$0.67$, $1.01$] & . & $1.19$ &  [$0.81$, $1.74$] &  & $1.03$ &  [$0.82$, $1.30$] &  & $1.42$ &  [$1.07$, $1.90$] & * \\ 
\midrule
\multicolumn{13}{l}{Age} \\ 
\midrule
30-44 & $1.08$ &  [$0.78$, $1.49$] &  & $0.87$ &  [$0.43$, $1.75$] &  & $0.79$ &  [$0.55$, $1.13$] &  & $0.94$ &  [$0.59$, $1.49$] &  \\ 
45-59 & $1.13$ &  [$0.80$, $1.60$] &  & $0.95$ &  [$0.46$, $1.95$] &  & $0.92$ &  [$0.62$, $1.35$] &  & $0.76$ &  [$0.48$, $1.22$] &  \\ 
60+ & $0.96$ &  [$0.69$, $1.34$] &  & $0.70$ &  [$0.35$, $1.40$] &  & $1.01$ &  [$0.70$, $1.45$] &  & $0.73$ &  [$0.46$, $1.15$] &  \\ 
\midrule
\multicolumn{13}{l}{Education} \\ 
\midrule
HS or equivalent & $0.77$ &  [$0.46$, $1.30$] &  & $0.50$ &  [$0.20$, $1.21$] &  & $0.81$ &  [$0.47$, $1.40$] &  & $1.22$ &  [$0.66$, $2.27$] &  \\ 
Some college & $0.91$ &  [$0.56$, $1.49$] &  & $0.74$ &  [$0.32$, $1.71$] &  & $0.63$ &  [$0.38$, $1.06$] & . & $1.17$ &  [$0.65$, $2.12$] &  \\ 
Bachelor & $0.79$ &  [$0.48$, $1.32$] &  & $1.12$ &  [$0.42$, $2.97$] &  & $0.53$ &  [$0.31$, $0.93$] & * & $1.82$ &  [$0.91$, $3.64$] & . \\ 
Post graduate & $0.84$ &  [$0.49$, $1.44$] &  & $1.16$ &  [$0.41$, $3.30$] &  & $0.51$ &  [$0.28$, $0.91$] & * & $2.18$ &  [$1.04$, $4.57$] & * \\ 
\midrule
\multicolumn{13}{l}{Income} \\ 
\midrule
\$30k - \$60k & $1.19$ &  [$0.87$, $1.63$] &  & $1.25$ &  [$0.77$, $2.05$] &  & $1.00$ &  [$0.71$, $1.41$] &  & $1.22$ &  [$0.82$, $1.82$] &  \\ 
\$60k - \$100k & $1.23$ &  [$0.90$, $1.68$] &  & $2.14$ &  [$1.18$, $3.86$] & * & $0.87$ &  [$0.60$, $1.27$] &  & $1.88$ &  [$1.24$, $2.87$] & ** \\ 
more than \$100k & $1.36$ &  [$0.99$, $1.88$] & . & $1.57$ &  [$0.86$, $2.86$] &  & $0.86$ &  [$0.59$, $1.25$] &  & $2.18$ &  [$1.34$, $3.56$] & ** \\ 
\bottomrule
\end{longtable}
\begin{minipage}{\linewidth}
Signif. codes:  0 `***' 0.001 `**' 0.01 `*' 0.05 `.' 0.1 ` ' 1\\
\end{minipage}

\elandscape

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

Through this work, we have tested the use of probability-based survey
panels to ask perception questions; within a large survey covering a
variety of topics, and with a limited number of questions for our
specific tasks, we are able to measure viewer perception and produce
results consistent with prior studies. Testing data visualization design
structures and viewer behavior on a nationally-representative sample is
of broad interest and applicability to the scientific communication
community, and this study demonstrates a framework under which to
complete further work in this area.

With our survey framework, we have shown that we can rank structural
variations on design and tasks by their relative accuracy; we identified
significantly different rates of respondent accuracy under different
designs presenting the same data. The size of our sample aids in
identifying this signal.

Further, we can also study respondent behavior and how viewers interact
with the chart. Paradata on viewer interaction with the chart (e.g.,
zooming and time spent on each task) can be collected as well as asking
directly asking respondents questions about their certainty. Not only
can we collect this information, but we can also identify significant
differences in this behavior across different designs. Understanding
viewer interaction and engagement is a key component to designing
effective data visualizations. If a particular design leads viewers to
zoom in more to investigate it when determining a response, we might
consider that design to be more frustrating to viewers `in the wild'
when viewing it outside of the context of our tests.
\kr{I don't love the last sentence but I think we can clean it up}.

Perhaps most salient among our results are the patterns in response
behavior across demographic groups. The large size of our sample, paired
with the probability-based survey panel approach, afford us the ability
to dive into response patterns among demographics across tasks and
identify significant differences among them. In particular, the
differences across educational attainment groups and income groups
underscore the importance of utilizing a representative population when
testing perception; prior results in this area may be absorbing bias in
responses due to the larger representation of individuals with higher
education levels among those study respondents.

While our results are limited to mapping in a stacked bar chart with a
specific topic, they demonstrate among them a wealth of insights about
how respondents perceive and interact with data visualizations. More
expansive work on this topic should be completed to understand how the
general public interacts with and understands charts, including
expanding to a wider set of structural variations, more aesthetic design
variations, and considering differences in respondent behavior across
different data topic areas.

\hh{The actionable insights are:}

\begin{itemize}
\tightlist
\item
  make sure that the main comparison in a plot is lined up -- in an
  interactive setting it would be beneficial to have the possibility of
  re-ordering the categories to help with comparisons.
\item
  offering zoom-ins help with certainty, even if it does not help with
  perceptual accuracy
\end{itemize}

\hypertarget{supplementary-material}{%
\section*{Supplementary Material}\label{supplementary-material}}
\addcontentsline{toc}{section}{Supplementary Material}

\begin{itemize}
\tightlist
\item
  \textbf{Participant Data (Linear):} Link to csv file with the data.
\item
  \textbf{Data Analysis Code:} Link to an html document with annotated
  code chunks.
\end{itemize}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-borgoCrowdsourcingInformationVisualization2017}{}}%
Borgo, Rita, Bongshin Lee, Benjamin Bach, Sara Fabrikant, Radu Jianu,
Andreas Kerren, Stephen Kobourov, et al. 2017. {``Crowdsourcing for
{Information Visualization}: {Promises} and {Pitfalls}.''} In
\emph{Evaluation in the {Crowd}. {Crowdsourcing} and {Human-Centered
Experiments}}, edited by Daniel Archambault, Helen Purchase, and Tobias
Hoßfeld, 96--138. Lecture {Notes} in {Computer Science}. {Springer
International Publishing}.

\leavevmode\vadjust pre{\hypertarget{ref-clevelandGraphicalPerceptionTheory1984}{}}%
Cleveland, William S., and Robert McGill. 1984. {``Graphical Perception:
{Theory}, Experimentation, and Application to the Development of
Graphical Methods.''} \emph{Journal of the American Statistical
Association} 79 (387): 531--54.
\url{https://doi.org/10.1080/01621459.1984.10478080}.

\leavevmode\vadjust pre{\hypertarget{ref-dennis2019technical}{}}%
Dennis, J Michael. 2019, updated 2022. {``Technical Overview of the
{AmeriSpeak} Panel {NORC}'s Probability-Based Household Panel.''}
\emph{NORC at the University of Chicago}, 2019, updated 2022.

\leavevmode\vadjust pre{\hypertarget{ref-heerCrowdsourcingGraphicalPerception2010}{}}%
Heer, J., and M. Bostock. 2010. {``Crowdsourcing Graphical Perception:
{Using} Mechanical {Turk} to Assess Visualization Design.''} In
\emph{Conference on {Human Factors} in {Computing Systems} -
{Proceedings}}, 1:203--12. {ACM}.
\url{https://doi.org/10.1145/1753326.1753357}.

\leavevmode\vadjust pre{\hypertarget{ref-kishSurveySampling1965}{}}%
Kish, Leslie. 1965. \emph{Survey {Sampling}}. {Wiley}.

\leavevmode\vadjust pre{\hypertarget{ref-luModelingJustNoticeable2022}{}}%
Lu, Min, Joel Lanir, Chufeng Wang, Yucong Yao, Wen Zhang, Oliver
Deussen, and Hui Huang. 2022. {``Modeling {Just Noticeable Differences}
in {Charts}.''} \emph{IEEE Transactions on Visualization and Computer
Graphics} 28 (1): 718--26.
\url{https://doi.org/10.1109/TVCG.2021.3114874}.

\leavevmode\vadjust pre{\hypertarget{ref-lumleyAnalysisComplexSurvey2004}{}}%
Lumley, Thomas. 2004. {``Analysis of Complex Survey Samples.''}
\emph{Journal of Statistical Software} 9 (1): 1--19.

\leavevmode\vadjust pre{\hypertarget{ref-lumleyComplexSurveysGuide2010}{}}%
---------. 2010. \emph{Complex Surveys: {A} Guide to Analysis Using {R}:
{A} Guide to Analysis Using {R}}. {John Wiley and Sons}.

\leavevmode\vadjust pre{\hypertarget{ref-survey}{}}%
---------. 2020. {``Survey: Analysis of Complex Survey Samples.''}

\leavevmode\vadjust pre{\hypertarget{ref-mackinlayAutomatingDesignGraphical1986}{}}%
Mackinlay, Jock. 1986. {``Automating the Design of Graphical
Presentations of Relational Information.''} \emph{ACM Transactions on
Graphics} 5 (2): 110--41. \url{https://doi.org/10.1145/22949.22950}.

\leavevmode\vadjust pre{\hypertarget{ref-omuircheartaighCombiningSamplesVs2002}{}}%
O'Muircheartaigh, Colm, and Steven Pedlow. 2002. {``Combining {Samples
Vs}. {Cumulating Cases}: {A Comparison} of {Two Weighting Strategies} in
{NLSY97}.''} In \emph{{ASA Proceedings} of the {Joint Statistical
Meetings}}, 2557--62.
\url{http://www.asasrms.org/Proceedings/y2002/Files/JSM2002-001082.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-piephoAlgorithmLetterBasedRepresentation2004}{}}%
Piepho, Hans-Peter. 2004. {``An {Algorithm} for a {Letter-Based
Representation} of {All-Pairwise Comparisons}.''} \emph{Journal of
Computational and Graphical Statistics} 13 (2): 456--66.
\url{https://doi.org/10.1198/1061860043515}.

\leavevmode\vadjust pre{\hypertarget{ref-RLanguage}{}}%
R Core Team. 2022. \emph{R: {A} Language and Environment for Statistical
Computing}. Manual. {Vienna, Austria}: {R Foundation for Statistical
Computing}. \url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-talbotFourExperimentsPerception2014}{}}%
Talbot, Justin, Vidya Setlur, and Anushka Anand. 2014. {``Four
{Experiments} on the {Perception} of {Bar Charts}.''} \emph{IEEE
Transactions on Visualization and Computer Graphics} 20 (12): 2152--60.
\url{https://doi.org/10.1109/TVCG.2014.2346320}.

\leavevmode\vadjust pre{\hypertarget{ref-zacksReadingBarGraphs1998a}{}}%
Zacks, Jeff, Ellen Levy, Barbara Tversky, and Diane J. Schiano. 1998.
{``Reading Bar Graphs: {Effects} of Extraneous Depth Cues and Graphical
Context.''} \emph{Journal of Experimental Psychology: Applied} 4 (2):
119--38. \url{https://doi.org/10.1037/1076-898X.4.2.119}.

\end{CSLReferences}

\end{document}