### 2023-07-31 SDSS-supplement/supplement.Rmd
adding supplement PDF

---
title: "Testing Charts: viewers' perceptual accuracy in surveys"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Supplementary Materials

Additional materials for this paper are shared in the following GitHub repository: [https://github.com/kiegan/testing-charts-jds](https://github.com/kiegan/testing-charts-jds). The repository contains all data used in analyses, the original abstract submission file for SDSS 2023, and the code used to generate all figures and tables in the submission document.